### 2022-02-03 analysis/paper/002-paper-m-original-models.Rmd

```{r m-original-models-prep}
## extract the peak heights
d_res_path <- "../data/derived_data/d_peaks"
dir.create(d_res_path)

# extract scaled normalized peak heights from MIRS
res <- 
  hklmirs::hkl_get_normalized_peak_heights(
    x = d_flat,
    export = d_res_path,
    verbose = FALSE,
    make_plots = FALSE,
    do_scale = TRUE
  )

# extract information on scaling for prediction for new datasets
res_scale <- 
  tibble::tibble(
    name = colnames(res),
    center = purrr::map_dbl(res, attr, which = "scaled:center"),
    scale = purrr::map_dbl(res, attr, which = "scaled:scale")
  )

# add peak info to d
d <- dplyr::bind_cols(d, res %>% purrr::map_df(as.numeric))
```

```{r m-original-models-models}
# compute the regression models
m1.1 <- lm(hol ~ carb, 
         data = d %>% dplyr::filter(sample_type != "old magazines"))
m2.1 <- lm(kl ~ arom15arom16, 
         data = d %>% dplyr::filter(sample_type != "office paper"))
```

```{r m-original-models-or-bayesian}
## compute the original regression models as Bayesian models

# define priors
rstanarm_prior <- rstanarm::normal(0, 2.5, autoscale = TRUE)
rstanarm_prior_aux <- rstanarm::exponential(1, autoscale = TRUE)

# compute models
m1.2 <- rstanarm::stan_glm(hol ~ carb, 
                          data = d %>% dplyr::filter(sample_type != "old magazines"),
                          family = gaussian,
                          chains = chains,
                          iter = iter,
                          warmup = warmup,
                          seed = seed,
                          prior = rstanarm_prior,
                          prior_intercept = rstanarm_prior,
                          prior_aux = rstanarm_prior_aux)
m2.2 <- rstanarm::stan_glm(kl ~ arom15arom16, 
                          data = d %>% dplyr::filter(sample_type != "office paper"),
                          family = gaussian,
                          chains = chains,
                          iter = iter,
                          warmup = warmup,
                          seed = seed,
                          prior = rstanarm_prior,
                          prior_intercept = rstanarm_prior,
                          prior_aux = rstanarm_prior_aux)

# compute loos
m1.2_loo <- loo(m1.2, 
                k_threshold = k_threshold)
m2.2_loo <- loo(m2.2, 
                k_threshold = k_threshold)
```