### 2022-02-03 analysis/paper/003-paper-m-gaussian-beta.Rmd

```{r m-gaussian-beta-1}
## beta regression models

# define priors
brms_beta_priors <-
  c(
    brms::prior_string("normal(0, 2.5)", class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

# compute models
m1.3 <- brms::brm(hol ~ carb, 
                  data = d %>% dplyr::filter(sample_type != "old magazines"), 
                  family = Beta(link = "logit", link_phi = "log"),
                  prior = brms_beta_priors,
                  chains = chains,
                  seed = seed,
                  iter = iter,
                  warmup = warmup)

m2.3 <- brms::brm(kl ~ arom15arom16, 
                  data = d %>% dplyr::filter(sample_type != "office paper"), 
                  family = Beta(link = "logit", link_phi = "log"),
                  prior = brms_beta_priors,
                  chains = chains,
                  seed = seed,
                  iter = iter,
                  warmup = warmup)

# compute loos
m1.3_loo <- loo(m1.3, k_threshold = k_threshold)
m2.3_loo <- loo(m2.3, k_threshold = k_threshold)

# ---todo: brms and rstanarm yield very different uncertainty estimates. brms has smaller prediction intervals. The reason seems to be the different priors for phi (normal(0, 2.5) for rstanarm; gamma(0.01, 0.01) for brms).
```

```{r m-gaussian-beta-2}
## predictions for peat samples

# predict Klason lignin and holocellulose contents using the original model
d_peat_res_path <- "../data/derived_data/d_peat_peaks"
dir.create(d_peat_res_path)

d_peat <-
  d_peat %>%
  irpeat::irp_content_klh_hodgkins(export = d_peat_res_path, 
                                   verbose = FALSE, 
                                   make_plots = FALSE)

# import the extracted peaks and compute arom15arom16
d_peat_res <- read.csv(paste0(d_peat_res_path, "/Heights_Norm.Corr.csv"), 
                       header = TRUE, 
                       row.names = 1) %>%
  dplyr::mutate(arom15arom16 = arom15 + arom16) %>%
  purrr::map2_df(seq_len(nrow(res_scale)), function(x, i) {
    scale(x, 
          center = res_scale$center[[i]], 
          scale = res_scale$scale[[i]])
  })

# compute predictions
d_peat_pred <- 
  dplyr::bind_rows(
    m1.1 %>%
      predict(newdata = d_peat_res %>% dplyr::mutate(carb = as.numeric(carb)),
              interval = "prediction",
              level = 0.95) %>%
      as.data.frame() %>%
      dplyr::mutate(
        model = "Gaussian, non-Bayesian",
        variable_y = "Holocellulose"
      ),
    m1.2 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Gaussian, Bayesian",
        variable_y = "Holocellulose"
      ),
    m2.1 %>%
      predict(newdata = d_peat_res %>% dplyr::mutate(arom15arom16 = as.numeric(arom15arom16)),
              interval = "prediction",
              level = 0.95) %>%
      as.data.frame() %>%
      dplyr::mutate(
        model = "Gaussian, non-Bayesian",
        variable_y = "Klason lignin"
      ),
    m2.2 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Gaussian, Bayesian",
        variable_y = "Klason lignin"
      ),
    m1.3 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Beta",
        variable_y = "Holocellulose"
      ),
    m2.3 %>%
      rstanarm::posterior_predict(newdata = d_peat_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Beta",
        variable_y = "Klason lignin"
      )
  ) %>%
  dplyr::bind_cols(
    purrr::map_df(1:6, function(x) {
      tibble::tibble(
        sample_id = d_peat$sample_id, 
        core_label = d_peat$core_label,
        sampling_longitude = d_peat$sampling_longitude, 
        sampling_latitude = d_peat$sampling_latitude, 
        sampling_altitude = d_peat$sampling_altitude, 
        sample_depth_lower = d_peat$sample_depth_lower, 
        sample_depth_upper = d_peat$sample_depth_upper
      )
    })
  )

# add fitted values from Gaussian model as separate variable
d_peat_pred <-
  dplyr::left_join(d_peat_pred,
                   d_peat_pred %>% 
                     dplyr::filter(model == "Gaussian, Bayesian") %>%
                     dplyr::select(variable_y, sample_id, fit) %>%
                     dplyr::rename(fit_gaussian = "fit"),
                   by = c("variable_y", "sample_id"))
```

```{r m-gaussian-beta-3}
## predictions for vegetation samples

## create data sets for predictions
d_veg_res_path <- "../data/derived_data/d_veg_peaks"
dir.create(d_veg_res_path)
d_veg_res <-
  irpeat::irp_content_klh_hodgkins(d_veg,
                                   export = d_veg_res_path,
                                   verbose = FALSE,
                                   make_plots = FALSE)

# import the extracted features and compute missing variables
d_veg_res <- read.csv(paste0(d_veg_res_path, "/Heights_Norm.Corr.csv"), 
                       header = TRUE, 
                       row.names = 1) %>%
  dplyr::mutate(arom15arom16 = arom15 + arom16) %>%
  purrr::map2_df(seq_len(nrow(res_scale)), function(x, i) {
    scale(x, 
          center = res_scale$center[[i]], 
          scale = res_scale$scale[[i]])
  })

# compute predictions
d_veg_pred <- 
  dplyr::bind_rows(
    m1.1 %>%
      predict(newdata = d_veg_res %>% dplyr::mutate(carb = as.numeric(carb)),
              interval = "prediction",
              level = 0.95) %>%
      as.data.frame() %>%
      dplyr::mutate(
        model = "Gaussian, non-Bayesian",
        variable_y = "Holocellulose"
      ),
    m1.2 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Gaussian, Bayesian",
        variable_y = "Holocellulose"
      ),
    m2.1 %>%
      predict(newdata = d_veg_res %>% dplyr::mutate(arom15arom16 = as.numeric(arom15arom16)),
              interval = "prediction",
              level = 0.95) %>%
      as.data.frame() %>%
      dplyr::mutate(
        model = "Gaussian, non-Bayesian",
        variable_y = "Klason lignin"
      ),
    m2.2 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Gaussian, Bayesian",
        variable_y = "Klason lignin"
      ),
    m1.3 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Beta",
        variable_y = "Holocellulose"
      ),
    m2.3 %>%
      rstanarm::posterior_predict(newdata = d_veg_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        model = "Beta",
        variable_y = "Klason lignin"
      )
  ) %>%
  dplyr::bind_cols(
    purrr::map_df(1:6, function(x) {
      tibble::tibble(
        sample_id = d_veg$sample_id, 
        site_name = d_veg$site_name,
        species_name = d_veg$species_name,
        plant_part = d_veg$plant_part
      )
    })
  )

# add fitted values from Gaussian model as separate variable
d_veg_pred <-
  dplyr::left_join(d_veg_pred,
                   d_veg_pred %>% 
                     dplyr::filter(model == "Gaussian, Bayesian") %>%
                     dplyr::select(variable_y, sample_id, fit) %>%
                     dplyr::rename(fit_gaussian = "fit"),
                   by = c("variable_y", "sample_id"))
```

```{r m-gaussian-beta-4}
### plot of depth profiles for the peat core data: comparison of beta regression and Gaussian regression models

p1 <-
  d_peat_pred %>%
  dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Holocellulose", 
       y = expression(Holocellulose~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse()

p2 <-
  d_peat_pred %>%
  dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
  ggplot(aes(y = fit, ymin = lwr, ymax = upr, 
             x = sample_depth_lower, 
             colour = model, 
             fill = model)) +
  geom_ribbon(colour = NA, 
              alpha = 0.2) +
  geom_path() +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  coord_flip() +
  scale_y_continuous(labels = function(x) x * 100) +
  labs(title = "Klason lignin", 
       y = expression(Klason~lignin~content~"[mass-%]"),
       x = "Lower layer boundary depth [cm]") +
  facet_wrap(~ core_label, 
             scales = "free_y", 
             ncol = 2) +
  scale_x_reverse()

# combine plots
p_gaussian_beta_depth_profile <-
  p1 + p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

```{r m-gaussian-beta-5}
### prediction across the whole possible range from 0 to 100 mass-%: comparison of beta regression and Gaussian regression models

# prediction range
d_range_res <-
  tibble::tibble(
    variable_y = c("holocellulose", "Klason lignin"),
    from = rep(-13, 2),
    to = from * (-1)
  )

# sequence of values across the prediction range
d_range_res <- 
  tibble::tibble(
    carb = 
      seq(from = d_range_res$from[d_range_res$variable_y == "holocellulose"], 
          to = d_range_res$to[d_range_res$variable_y == "holocellulose"], 
          length.out = 2 * 1e2),
    arom15arom16 = 
      seq(from = d_range_res$from[d_range_res$variable_y == "Klason lignin"], 
          to = d_range_res$to[d_range_res$variable_y == "Klason lignin"], 
          length.out = 2 * 1e2)
  )

# predictions
d_range_pred <- 
  dplyr::bind_rows(
    m1.1 %>%
      predict(newdata = d_range_res %>% dplyr::mutate(carb = as.numeric(carb)),
              interval = "prediction",
              level = 0.95) %>%
      as.data.frame() %>%
      dplyr::mutate(
        model = "Gaussian, non-Bayesian",
        variable_y = "Holocellulose"
      ),
    m1.2 %>%
      rstanarm::posterior_predict(newdata = d_range_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        iter = seq_len(nrow(.)),
        model = "Gaussian, Bayesian",
        variable_y = "Holocellulose"
      ),
    m2.1 %>%
      predict(newdata = d_range_res %>% dplyr::mutate(arom15arom16 = as.numeric(arom15arom16)),
              interval = "prediction",
              level = 0.95) %>%
      as.data.frame() %>%
      dplyr::mutate(
        model = "Gaussian, non-Bayesian",
        variable_y = "Klason lignin"
      ),
    m2.2 %>%
      rstanarm::posterior_predict(newdata = d_range_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        iter = seq_len(nrow(.)),
        model = "Gaussian, Bayesian",
        variable_y = "Klason lignin"
      ),
    m1.3 %>%
      brms::posterior_predict(newdata = d_range_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        iter = seq_len(nrow(.)),
        model = "Beta",
        variable_y = "Holocellulose"
      ),
    m2.3 %>%
      brms::posterior_predict(newdata = d_range_res) %>%
      as.data.frame() %>%
      purrr::map_df(function(x) {
        tibble::tibble(
          fit = median(x),
          lwr = quantile(x, prob = 0.05),
          upr = quantile(x, prob = 0.95)
        )
      }) %>%
      dplyr::mutate(
        iter = seq_len(nrow(.)),
        model = "Beta",
        variable_y = "Klason lignin"
      )
  ) %>%
  dplyr::bind_cols(
    purrr::map_df(1:6, function(x) {
      d_range_res
    })
  )

# add fitted values from Gaussian model as separate variable
d_range_pred <-
  dplyr::left_join(d_range_pred,
                   d_range_pred %>% 
                     dplyr::filter(model == "Gaussian, Bayesian") %>%
                     dplyr::select(variable_y, iter, fit) %>%
                     dplyr::rename(fit_gaussian = "fit"),
                   by = c("variable_y", "iter"))

# detect fitted values where prediction intervals produce unrealistic predictions for the Gaussian models
d_range_pred_limits <-
  d_range_pred %>%
  dplyr::filter(model == "Gaussian, Bayesian") %>%
  dplyr::group_by(model, variable_y) %>%
  dplyr::summarise(index_fit_lwr = which(lwr > 0)[[1]],
                   index_fit_upr = which(upr > 1)[[1]],
                   fit_lwr = fit[index_fit_lwr],
                   fit_upr = fit[index_fit_upr],
                   lwr_lwr = lwr[index_fit_lwr],
                   upr_upr = upr[index_fit_upr],
                   fit_gaussian_lwr = fit_gaussian[index_fit_lwr],
                   fit_gaussian_upr = fit_gaussian[index_fit_upr],
                   .groups = "keep")

d_range_pred_limits <-
  dplyr::bind_cols(
    d_range_pred_limits %>%
      tidyr::pivot_longer(cols = dplyr::all_of(c("fit_lwr", "fit_upr")), 
                          values_to = "fit", 
                          names_to = "variable_limit"),
    d_range_pred_limits %>%
      tidyr::pivot_longer(cols = dplyr::all_of(c("fit_gaussian_lwr", "fit_gaussian_upr")), 
                          values_to = "fit_gaussian", 
                          names_to = "variable_gaussian_limit") %>%
      dplyr::ungroup() %>%
      dplyr::select(fit_gaussian)
  )
```

```{r m-gaussian-beta-6}
# plot
p1 <-
  d_range_pred %>%
  dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian") %>%
  ggplot(aes(x = fit_gaussian, y = fit, colour = model, fill = model)) +
  geom_hline(yintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_vline(xintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, colour = NA) +
  geom_path() +
  geom_point(data = d_peat_pred %>% dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian"), 
             mapping = aes(x = fit_gaussian, y = fit)) +
  geom_point(data = d_veg_pred %>% dplyr::filter(variable_y == "Holocellulose" & model != "Gaussian, non-Bayesian"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 4) +
  geom_point(data = d_range_pred_limits %>% dplyr::filter(variable_y == "Holocellulose"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 21, fill = "white", colour = "black") +
  labs(title = "Holocellulose", 
       x = expression(Predicted~values~"(Gaussian model)"),
       y = expression(Predicted~values)) +
  coord_cartesian(ylim = c(0, 1), xlim = c(0, 1)) +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100)

p2 <-
  d_range_pred %>%
  dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian") %>%
  ggplot(aes(x = fit_gaussian, y = fit, colour = model, fill = model)) +
  geom_hline(yintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_vline(xintercept = c(0, 1), linetype = 2, colour = "grey") +
  geom_ribbon(aes(ymin = lwr, ymax = upr), alpha = 0.2, colour = NA) +
  geom_path() +
  geom_point(data = d_peat_pred %>% dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian"), 
             mapping = aes(x = fit_gaussian, y = fit)) +
  geom_point(data = d_veg_pred %>% dplyr::filter(variable_y == "Klason lignin" & model != "Gaussian, non-Bayesian"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 4) +
  geom_point(data = d_range_pred_limits %>% dplyr::filter(variable_y == "Klason lignin"), 
             mapping = aes(x = fit_gaussian, y = fit), shape = 21, fill = "white", colour = "black") +
  labs(title = "Klason lignin", 
       x = expression(Predicted~values~"(Gaussian model)"),
       y = expression(Predicted~values)) +
  coord_cartesian(ylim = c(0, 1), xlim = c(0, 1)) +
  scale_x_continuous(labels = function(x) x * 100) +
  scale_y_continuous(labels = function(x) x * 100)

# combine plots
p_gaussian_beta_range <-
  p1 + p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```