### 2022-02-03 analysis/paper/005-paper-m-minerals.Rmd

```{r m-minerals-1}
## Holocellulose

m1.12 <- 
  brms::brm(hol ~ carb + arom15 + arom16 + acids + aliph28 + aliph29 + trough16 + trough28, 
            data = d, 
            family = Beta(link = "logit", link_phi = "log"),
            prior = brms_beta_priors1,
            chains = chains,
            iter = iter,
            warmup = warmup,
            seed = seed)

brms_beta_priors2 <-
  c(
    brms::prior_string(paste0("horseshoe(df = 1, scale_global = ", prior_hs_tau0[[4]], ", df_global = 1, autoscale = TRUE)"),
                       class = "b"),
    brms::prior_string("normal(0, 2.5)", class = "Intercept"),
    brms::prior_string("gamma(0.01, 0.01)", class = "phi")
  )

m1.13 <- 
  brms::brm(
    hol ~ ., 
    data = 
      d_underfitting[[4]] %>% 
      dplyr::select(-kl),
    family = Beta(link = "logit", link_phi = "log"),
    prior = brms_beta_priors2,
    cores = chains,
    chains = chains,
    warmup = warmup,
    iter = iter,
    seed = seed,
    control = list(adapt_delta = 0.99, max_treedepth = 15) # with default adapt_delta (0.8): 22 divergent transitions; adapt_delta = 0.99: There were 2000 transitions after warmup that exceeded the maximum treedepth. Increase max_treedepth above 10
  ) 
```

```{r m-minerals-2}
# PSIS-LOO
m1.12_loo <- loo(m1.12, reloo = TRUE, cores = chains)
m1.13_loo <- loo(m1.13, reloo = TRUE, cores = chains)

loo_comparison_mineral <- loo::loo_compare(m1.12_loo, m1.13_loo)
```

```{r m-minerals-3}
# plot of fitted values against measured values
d_minerals_pred <-
  tibble::tibble(
    fit = c(
      posterior_predict(m1.2, newdata = d) %>%
        apply(2, median),
      posterior_predict(m1.10, newdata = d_underfitting[[4]]) %>%
        apply(2, median),
      posterior_predict(m1.6, newdata = d) %>%
        apply(2, median),
      posterior_predict(m1.12) %>%
        apply(2, median),
      posterior_predict(m1.13) %>%
        apply(2, median)
    ),
    model = 
      rep(c("Original", "Binned (no minerals)", "Peaks (no minerals)", "Peaks", "Binned"), each = nrow(d)) %>%
      factor(levels = c("Original", "Peaks (no minerals)", "Binned (no minerals)", "Peaks", "Binned"))
  ) %>%
  dplyr::bind_cols(
    purrr::map_df(1:5, function(x) {
      d %>% 
        dplyr::select(sample_type, hol) %>%
        dplyr::mutate(mineral = ifelse(sample_type == "old magazines", TRUE, FALSE))
    })
  )

# plot
p_y_yhat_minerals <-
  d_minerals_pred %>%
  ggplot(aes(x = fit * 100, y = hol * 100, colour = mineral)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey") +
  geom_point() +
  coord_fixed() +
  facet_wrap( ~ model, nrow = 1) +
  labs(x = "Fitted",
       y = "Measured") +
  guides(colour = guide_legend(title = "Mineral rich samples")) +
  theme(legend.position = "bottom")
```

```{r m-p-coefficients}
## plot to compare coefficients of the best models

# Holocellulose
m1.10_pars <- 
  m1.10 %>%
  as.data.frame() %>%
  dplyr::select(dplyr::contains("b_V"))

m1.10_pars <- 
  m1.10_pars %>%
  purrr::map_df(function(x) {
    res <- quantile(x, c(0.5, 0.05, 0.95))
    tibble::tibble(
      median = res[[1]],
      lwr = res[[2]],
      upr = res[[3]]
    )
  }) %>%
  dplyr::mutate(
    par_name = colnames(m1.10_pars),
    wn = d_underfitting_flat[[4]]$x,
    variable_y = "Holocellulose (no minerals)"
  )

m1.13_pars <- 
  m1.13 %>%
  as.data.frame() %>%
  dplyr::select(dplyr::contains("b_V"))

m1.13_pars <- 
  m1.13_pars %>%
  purrr::map_df(function(x) {
    res <- quantile(x, c(0.5, 0.05, 0.95))
    tibble::tibble(
      median = res[[1]],
      lwr = res[[2]],
      upr = res[[3]]
    )
  }) %>%
  dplyr::mutate(
    par_name = colnames(m1.13_pars),
    wn = d_underfitting_flat[[4]]$x,
    variable_y = "Holocellulose (with minerals)"
  )

# Klason lignin
m2.7_pars <- 
  m2.7 %>%
  as.data.frame() %>%
  dplyr::select(dplyr::contains("b_V"))

m2.7_pars <- 
  m2.7_pars %>%
  purrr::map_df(function(x) {
    res <- quantile(x, c(0.5, 0.05, 0.95))
    tibble::tibble(
      median = res[[1]],
      lwr = res[[2]],
      upr = res[[3]]
    )
  }) %>%
  dplyr::mutate(
    par_name = colnames(m2.7_pars),
    wn = d_underfitting_flat[[1]]$x,
    variable_y = "Klason lignin"
  )

# combine
m_pars <-
  dplyr::bind_rows(
    m1.10_pars,
    m1.13_pars,
    m2.7_pars
  )

# plots
p_coefficients <-
  m_pars %>%
  ggplot(aes(x = wn, y = median)) +
  geom_ribbon(data = d_peat$spectra[[1]], aes(x = x, y = y * 200 + 0.4, ymax = y * 200 + 0.4, ymin = rep(0.4, length(y))), colour = NA, fill = "lightgrey", alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = 2, colour = "grey") +
  geom_errorbar(aes(ymin = lwr, ymax = upr), colour = "lightblue", width = 0) +
  geom_point(shape = 21, fill = "white", colour = "grey50") +
  facet_wrap(~ variable_y, scales = "free_x") +
  geom_text_repel(data = m_pars %>%
                    dplyr::mutate(label = ifelse(abs(median) >= 0.02, wn %>% round(0), "")),
                  aes(label = label),
                  segment.colour = "grey50") +
  labs(x = expression(Bin~wavenumber~"["*cm^{-1}*"]"),
       y = "Coefficient")
```

```{r m-model-validation}
d_models <-
  tibble::tibble(
    model = list(m1.2, m1.3, m1.5, m1.6, m1.7, m1.8, m1.9, m1.10, m1.11, m1.12, m1.13, 
                 m2.2, m2.3, m2.5, m2.6, m2.7, m2.8, m2.9, m2.10, m2.11),
    model_type = purrr::map_chr(model, function(x) class(x)[[1]]),
    variable_y = c(rep("Holocellulose", 11), rep("Klason lignin", 9)),
    n_divergent_transitions = 
      purrr::map2_dbl(model, model_type, function(x, y) {
        switch(y,
               "brmsfit" = sum(rstan::get_divergent_iterations(x$fit)),
               "stanreg" = sum(rstan::get_divergent_iterations(x$stanfit)),
               stop("Unknown model type.")
        )
      }
      )
  ) %>%
  dplyr::select(-model)
```