### 2022-03-03 analysis/paper/007-paper-m-prediction-differences.Rmd

```{r m-prediction-differences-1}
# create a data frame for plotting
d_pred_diff <-
  dplyr::bind_rows(
    dplyr::bind_rows(
    ## training data
    # Original Gaussian
    d %>%
      dplyr::mutate(
        hol_hat =
          m1.2 %>%
          posterior_predict(newdata = d) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        kl_hat = 
          m2.2 %>%
          posterior_predict(newdata = d) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Original Gaussian"
      ) %>%
      dplyr::select(-spectra),
    # Original Beta
    d %>%
      dplyr::mutate(
        hol_hat =
          m1.3 %>%
          posterior_predict(newdata = d) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        kl_hat = 
          m2.3 %>%
          posterior_predict(newdata = d) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Original Beta"
      ) %>%
      dplyr::select(-spectra),
    # Best all peaks
    d %>%
      dplyr::mutate(
        hol_hat =
          m1.6 %>%
          posterior_predict(newdata = d) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
         kl_hat = 
          m2.6 %>%
          posterior_predict(newdata = d) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Best all peaks"
      ) %>%
      dplyr::select(-spectra),
    # Best binned spectra
    d %>%
      dplyr::mutate(
        hol_hat =
          m1.8 %>%
          posterior_predict(newdata = d_underfitting[[2]]) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
         kl_hat = 
          m2.8 %>%
          posterior_predict(newdata = d_underfitting[[2]]) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Best binned spectra"
      ) %>%
      dplyr::select(-spectra)
    ) %>%
      dplyr::mutate(data_type = "Training"),

    dplyr::bind_rows(
    ## peat data
    # Original Gaussian
    d_peat %>%
      dplyr::mutate(
        hol_hat =
          m1.2 %>%
          posterior_predict(newdata = d_peat_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        kl_hat = 
          m2.2 %>%
          posterior_predict(newdata = d_peat_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Original Gaussian"
      ) %>%
      dplyr::select(-spectra),
    # Original Beta
    d_peat %>%
      dplyr::mutate(
        hol_hat =
          m1.3 %>%
          posterior_predict(newdata = d_peat_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        kl_hat = 
          m2.3 %>%
          posterior_predict(newdata = d_peat_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Original Beta"
      ) %>%
      dplyr::select(-spectra),
    # Best all peaks
    d_peat %>%
      dplyr::mutate(
        hol_hat =
          m1.6 %>%
          posterior_predict(newdata = d_peat_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
         kl_hat = 
          m2.6 %>%
          posterior_predict(newdata = d_peat_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Best all peaks"
      ) %>%
      dplyr::select(-spectra),
    # Best binned spectra
    d_peat %>%
      dplyr::mutate(
        hol_hat =
          m1.8 %>%
          posterior_predict(newdata = d_peat_bin20) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
         kl_hat = 
          m2.8 %>%
          posterior_predict(newdata = d_peat_bin20) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Best binned spectra"
      ) %>%
      dplyr::select(-spectra)
    ) %>%
      dplyr::mutate(data_type = "Peat") %>%
      dplyr::select(-dplyr::contains(c("holocellulose_hodgkins", "klason_lignin_hodgkins"))) %>%
      dplyr::bind_cols(purrr::map_df(1:4, function(x) d_peat_res)),

    dplyr::bind_rows(
    ## vegetation data
    # Original Gaussian
    d_veg %>%
      dplyr::mutate(
        hol_hat =
          m1.2 %>%
          posterior_predict(newdata = d_veg_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        kl_hat = 
          m2.2 %>%
          posterior_predict(newdata = d_veg_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Original Gaussian"
      ) %>%
      dplyr::select(-spectra),
    # Original Beta
    d_veg %>%
      dplyr::mutate(
        hol_hat =
          m1.3 %>%
          posterior_predict(newdata = d_veg_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        kl_hat = 
          m2.3 %>%
          posterior_predict(newdata = d_veg_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Original Beta"
      ) %>%
      dplyr::select(-spectra),
    # Best all peaks
    d_veg %>%
      dplyr::mutate(
        hol_hat =
          m1.6 %>%
          posterior_predict(newdata = d_veg_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
         kl_hat = 
          m2.6 %>%
          posterior_predict(newdata = d_veg_res) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Best all peaks"
      ) %>%
      dplyr::select(-spectra),
    # Best binned spectra
    d_veg %>%
      dplyr::mutate(
        hol_hat =
          m1.8 %>%
          posterior_predict(newdata = d_veg_bin20) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
         kl_hat = 
          m2.8 %>%
          posterior_predict(newdata = d_veg_bin20) %>%
          as.data.frame() %>%
          purrr::map_dbl(median),
        model = "Best binned spectra"
      ) %>%
      dplyr::select(-spectra)
    ) %>%
      dplyr::mutate(data_type = "Vegetation") %>%
      dplyr::select(-dplyr::contains(c("holocellulose_hodgkins", "klason_lignin_hodgkins"))) %>%
      dplyr::bind_cols(purrr::map_df(1:4, function(x) d_veg_res))
  )

# add additional variables
d_pred_diff <-
  d_pred_diff %>%
  dplyr::mutate(
    hol_residuals = hol - hol_hat,
    kl_residuals = kl - kl_hat,
    hol_hat_or = 
      purrr::map(unique(d_pred_diff$data_type), function(x) {
        d_pred_diff %>% 
          dplyr::filter(model == "Original Gaussian" & data_type == x) %>%
          dplyr::pull(hol_hat) %>%
          rep(4)
      }) %>% 
      unlist(),
    kl_hat_or = 
      purrr::map(unique(d_pred_diff$data_type), function(x) {
        d_pred_diff %>% 
          dplyr::filter(model == "Original Gaussian" & data_type == x) %>%
          dplyr::pull(kl_hat) %>%
          rep(4)
      }) %>% 
      unlist(),
    arom15 = as.numeric(arom15) * res_scale$scale[res_scale$name == "arom15"] + res_scale$center[res_scale$name == "arom15"],
    arom16 = as.numeric(arom16) * res_scale$scale[res_scale$name == "arom16"] + res_scale$center[res_scale$name == "arom16"],
    arom15arom16 = arom15 + arom16,
    arom15arom16_rel = arom15/arom15arom16
  ) %>% 
  dplyr::mutate(
    model = factor(model, levels = c("Original Gaussian", "Original Beta", "Best all peaks", "Best binned spectra")),
    data_type = factor(data_type, levels = c("Training", "Peat", "Vegetation")),
    sample_type = factor(sample_type, levels = c("needles", "hardwood", "leaves and grasses", "softwood", "old corrugated cardboard", "old newsprint", "office paper", "old magazines"))
  )

# spectra
d_pred_diff_spectra <-
  ir::ir_new_ir(
    spectra = 
      dplyr::bind_rows(
        purrr::map_df(1:4, function(x) d %>% dplyr::select(spectra)),
        purrr::map_df(1:4, function(x) d_peat %>% dplyr::select(spectra)),
        purrr::map_df(1:4, function(x) d_veg %>% dplyr::select(spectra))
      ) %>% dplyr::pull(spectra),
    sample_id =
      d_pred_diff %>%
      nrow() %>%
      seq_len() %>%
      as.character()
  )

d_pred_diff_intensities <-
  d_pred_diff_spectra %>%
  ir::ir_get_intensity(wavenumber = c(3400, 1250, 1590)) %>%
  ir::ir_drop_spectra() %>%
  tidyr::unnest(cols = "intensity") %>%
  tidyr::pivot_wider(names_from = x, values_from = y, names_prefix = "intensity_")
```

```{r m-prediction-differences-2}
## plot of measured vs fitted values for the training data

# Holocellulose
p_index <- with(d_pred_diff, data_type == "Training" & 
                  sample_type != "old magazines" & 
                  model != "Original Beta")

p_pred_diff_train1_p1 <-
  d_pred_diff %>%
  dplyr::mutate(carb_rel = 
                  d_pred_diff_intensities %>%
                  dplyr::pull(intensity_1590) %>%
                  scale()
  ) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = hol_hat * 100, y = hol * 100)) +
  geom_point(aes(colour = scale(carb),
                 size = carb_rel)) +
  #geom_point(data = 
  #             d_pred_diff %>%
  #             dplyr::mutate(carb_rel = 
  #                
  #                  d_pred_diff_intensities %>%
  #                    dplyr::pull(intensity_1590)) %>%
  #             dplyr::filter(p_index & sample_id %in% d$sample_id[d$sample_type != "old magazines"][hol_sel]), 
  #           aes(x = hol_hat * 100, y = hol * 100,
  #               size = carb_rel), 
  #           fill = NA, shape = 21, colour = "grey50") +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(size = guide_legend(title = expression(Absorbance~at~1590~cm^{-1})),
         colour = guide_colourbar(title = "carb")) +
  labs(x = "Fitted values", y = "Measured values",
       title = "Holocellulose") +
  coord_fixed() +
  theme(legend.position = "bottom")

## Klason lignin

p_index <- with(d_pred_diff, data_type == "Training" & 
                  sample_type != "office paper" & 
                  model != "Original Beta")

p_pred_diff_train1_p2 <-
  d_pred_diff %>%
  dplyr::mutate(carb_rel = arom15arom16_rel) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = kl_hat * 100, y = kl * 100)) +
  geom_point(aes(colour = scale(arom15arom16),
                 size = carb_rel)) +
  #geom_point(data = 
  #             d_pred_diff %>%
  #             dplyr::mutate(carb_rel = arom15arom16_rel) %>%
  #             dplyr::filter(p_index & sample_id %in% d$sample_id[d$sample_type != "old magazines"][kl_sel]), 
  #           aes(x = kl_hat * 100, y = kl * 100,
  #               size = carb_rel), 
  #           fill = NA, shape = 21, colour = "grey50") +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(size = guide_legend(title = expression("arom15/(arom15arom16)")),
         colour = guide_colourbar(title = "arom15arom16")) +
  labs(x = "Fitted values", y = "Measured values",
       title = "Klason lignin") +
  coord_fixed() +
  theme(legend.position = "bottom")

# combine
p_pred_diff_train1 <-
  p_pred_diff_train1_p1 / p_pred_diff_train1_p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")")

### Start: test

hol_sel <- c(40, 18, 49)
hol_sel <- c(30, 39, 15)

d[d$sample_type != "old magazines", ][hol_sel, ]  %>%
  irpeat:::irp_content_klh_hodgkins_prepare() %>%
  hkl_hodgkins_ggplot() %>%
  wrap_plots()

d_pred_diff %>%
  dplyr::mutate(carb_rel = 
                  d_pred_diff_intensities %>%
                  dplyr::pull(intensity_1590)) %>%
  dplyr::filter(model == "Best binned spectra") %>%
  ggplot(aes(x = carb, y = (hol_hat - hol_hat_or) * 100)) +
  geom_point() +
  facet_grid(data_type ~ cut_number(carb_rel, 5))

p1_index <- with(d_pred_diff, (sample_type != "old magazines" | is.na(sample_type)) & 
                   model != "Original Beta")

p1 <-
  d_pred_diff %>%
  dplyr::mutate(carb_rel = 
                  d_pred_diff_intensities %>%
                  dplyr::pull(intensity_1590)) %>%
  dplyr::filter(p1_index) %>%
  ggplot(aes(x = hol_hat * 100, y = hol_hat_or * 100)) +
  geom_point(aes(colour = carb,
                 size = carb_rel,
                 shape = data_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(size = guide_legend(title = "carb/OH-peak")) +
  labs(x = "Fitted values", y = "Measured values") +
  coord_fixed()

hol_sel <- c(1, 11, 12, 6) # transect from underestimation to overestimation for the best binned model.

### End: test
```

```{r m-prediction-differences-3}
## plot of measured vs fitted values for the training data

# Holocellulose
p_index <- with(d_pred_diff, data_type == "Training" & 
                  sample_type != "old magazines" & 
                  model != "Original Beta")

p1 <-
  d_pred_diff %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = hol_hat * 100, 
             y = hol * 100)) +
  geom_point(aes(colour = sample_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(colour = guide_legend(title = "Sample type")) +
  labs(x = "Fitted values", y = "Measured values",
       title = "Holocellulose") +
  coord_fixed() +
  scale_colour_manual(values = scales::hue_pal()(length(levels(d_pred_diff$sample_type))),
                    labels = levels(d_pred_diff$sample_type), 
                    drop = FALSE)

## Klason lignin

p_index <- with(d_pred_diff, data_type == "Training" & 
                  sample_type != "office paper" & 
                  model != "Original Beta")

p2 <-
  d_pred_diff %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = kl_hat * 100, y = kl * 100)) +
  geom_point(aes(colour = sample_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(colour = guide_legend(title = "Sample type")) +
  labs(x = "Fitted values", y = "Measured values",
       title = "Klason lignin") +
  coord_fixed() +
  scale_colour_manual(values = scales::hue_pal()(length(levels(d_pred_diff$sample_type))),
                    labels = levels(d_pred_diff$sample_type), 
                    drop = FALSE)

# combine
p_pred_diff_train2 <-
  p1 / p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") +
  plot_layout(guides = "collect") &
  theme(legend.position = "right")

### Test start:

# in the plots above, one can see that the model makes different predictions for two OM classes: needles and leaves and grasses. How do these spectra differ? It seem that this is mainly th region around 1590
d %>% 
  dplyr::filter(sample_type %in% c("needles", "leaves and grasses")) %>%
  plot() + 
  geom_path(aes(x = x, y = y, colour = sample_type, group = measurement_id)) +
  geom_vline(xintercept = 1590)
```

```{r m-prediction-differences-4}
## plot of fitted (original model) vs fitted values for the training data

# Holocellulose
p_index <- with(d_pred_diff, 
                data_type == "Training" & 
                  sample_type != "old magazines" & 
                  model != "Original Beta")

p1 <-
  d_pred_diff %>%
  dplyr::mutate(carb_rel = 
                  d_pred_diff_intensities %>%
                  dplyr::pull(intensity_1590) %>%
                  scale()
  ) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = hol_hat * 100, y = hol_hat_or * 100)) +
  geom_point(aes(colour = scale(carb),
                 size = carb_rel)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(size = guide_legend(title = expression(Absorbance~at~1590~cm^{-1})),
         colour = guide_colourbar(title = "carb")) +
  labs(x = "Fitted values", y = "Fitted values (original model)",
       title = "Holocellulose") +
  coord_fixed() +
  theme(legend.position = "bottom")

## Klason lignin

p_index <- with(d_pred_diff, 
                data_type == "Training" & 
                  sample_type != "office paper" & 
                  model != "Original Beta")

p2 <-
  d_pred_diff %>%
  dplyr::mutate(carb_rel = scale(arom15arom16_rel)) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = kl_hat * 100, y = kl_hat_or * 100)) +
  geom_point(aes(colour = scale(arom15arom16),
                 size = carb_rel)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(size = guide_legend(title = expression("arom15/(arom15arom16)")),
         colour = guide_colourbar(title = "arom15arom16")) +
  labs(x = "Fitted values", y = "Fitted values (original model)",
       title = "Klason lignin") +
  coord_fixed() +
  theme(legend.position = "bottom")

# combine
p_pred_diff_train3 <-
  p1 / p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")")
```

```{r m-prediction-differences-5}
## plot of measured vs fitted values for the training data, peat data and vegetation data

# Holocellulose
p_index <- 
  with(
    d_pred_diff %>%
      dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
      dplyr::arrange(dplyr::desc(data_type)),
    (sample_type != "old magazines" | is.na(sample_type)) & 
      model != "Original Beta"
  )

p1 <-
  d_pred_diff %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  dplyr::arrange(dplyr::desc(data_type)) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = hol_hat * 100, y = hol_hat_or * 100)) +
  geom_point(aes(colour = data_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(colour = guide_legend(title = "Data set")) +
  labs(x = "Predicted values", y = "Predicted values (original model)",
       title = "Holocellulose") +
  coord_fixed() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

## Klason lignin

p_index <- 
  with(
    d_pred_diff %>%
      dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
      dplyr::arrange(dplyr::desc(data_type)),
    (sample_type != "office paper" | is.na(sample_type)) & 
      model != "Original Beta"
  )

p2 <-
  d_pred_diff %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  dplyr::arrange(dplyr::desc(data_type)) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = kl_hat * 100, y = kl_hat_or * 100)) +
  geom_point(aes(colour = data_type)) +
  geom_abline(intercept = 0, slope = 1, colour = "grey50") +
  facet_wrap( ~ model) +
  guides(colour = guide_legend(title = "Data set")) +
  labs(x = "Predicted values", y = "Predicted values (original model)",
       title = "Klason lignin") +
  coord_fixed() +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

# combine
p_pred_diff_all1 <-
  p1 / p2 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") + 
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")
```

```{r m-prediction-differences-6}
## for Klason lignin: residuals versus arom15arom16 conditional on arom15arom16_rel

p_index <- with(d_pred_diff, 
                data_type == "Training" & 
                  sample_type != "office paper" & 
                  model != "Original Beta")

p_pred_diff_kl_residuals <-
  d_pred_diff %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(x = arom15arom16 %>% scale(), y = kl_residuals)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, colour = "dimgrey") +
  geom_hline(yintercept = 0, colour = "grey50") +
  facet_grid(cut_number(arom15arom16_rel, n = 4) ~ model) +
  labs(x = "arom15arom16", y = "Residuals") + 
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "Standardized normalized arom15arom16 peak height", breaks = NULL, labels = NULL)) +
  scale_x_continuous(sec.axis = sec_axis(~ . , name = "Model", breaks = NULL, labels = NULL)) +
  theme(
    axis.line.y.right = element_line(color = "white"),
    axis.line.x.top = element_line(color = "white")
    )
```

```{r m-prediction-differences-7}
## for holocellulose: spectra across the gradient in predictions differences between the original model and the improved models

p_index <- with(d_pred_diff, 
                  (sample_type != "old magazines" | is.na(sample_type)) & 
                  model == "Best binned spectra")

p_pred_diff_hol_spectra_diff1 <-
  d_pred_diff_spectra %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  plot() +
  geom_path(aes(x = x, y = y, group = measurement_id), 
            colour = "white") +
  geom_path(aes(x = x, y = y, group = measurement_id, colour = hol_hat_or - hol_hat), 
            alpha = 0.3) +
  scale_colour_gradient2(low = scales::muted("blue"), 
                         mid = "white", 
                         high = scales::muted("red"))  +
  facet_wrap(~ data_type, ncol = 1) +
  guides(colour = guide_colorbar(title = "Prediction difference")) +
  theme(legend.position = "bottom") +
  labs(x = expression("Wavenumber ["*cm^{-1}*"]"), y = "Absorbance")

p_pred_diff_hol_spectra_diff2 <-
  d_pred_diff_spectra %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  plot() +
  geom_path(aes(x = x, y = y, group = measurement_id), 
            colour = "white") +
  geom_path(aes(x = x, y = y, group = measurement_id, colour = hol_hat_or - hol_hat), 
            alpha = 0.3) +
  scale_colour_gradient2(low = scales::muted("blue"), 
                         mid = "white", 
                         high = scales::muted("red"))  +
  facet_grid(cut_number(carb, 4) ~ data_type) +
  guides(colour = guide_colorbar(title = "Prediction difference")) +
  labs(x = expression("Wavenumber ["*cm^{-1}*"]"), y = "Absorbance") + 
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "Standardized normalized carb peak height", breaks = NULL, labels = NULL)) +
  scale_x_continuous(sec.axis = sec_axis(~ . , name = "Sample type", breaks = NULL, labels = NULL)) +
  theme(
    legend.position = "bottom",
    axis.line.y.right = element_line(color = "white"),
    axis.line.x.top = element_line(color = "white")
  )

## same for Klason lignin
p_index <- with(d_pred_diff, 
                  (sample_type != "office paper" | is.na(sample_type)) & 
                  model == "Best binned spectra")

p_pred_diff_kl_spectra_diff1 <-
  d_pred_diff_spectra %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, kl_hat_or, kl_hat, arom15arom16)
  ) %>%
  dplyr::filter(p_index) %>%
  plot() +
  geom_path(aes(x = x, y = y, group = measurement_id), 
            colour = "white") +
  geom_path(aes(x = x, y = y, group = measurement_id, colour = kl_hat_or - kl_hat), 
            alpha = 0.3) +
  scale_colour_gradient2(low = scales::muted("blue"), 
                         mid = "white", 
                         high = scales::muted("red"))  +
  facet_wrap(~ data_type, ncol = 1) +
  guides(colour = guide_colorbar(title = "Prediction difference")) +
  theme(legend.position = "bottom") +
  labs(x = expression("Wavenumber ["*cm^{-1}*"]"), y = "Absorbance")

p_pred_diff_kl_spectra_diff2 <-
  d_pred_diff_spectra %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, kl_hat_or, kl_hat, arom15arom16)
  ) %>%
  dplyr::filter(p_index) %>%
  plot() +
  geom_path(aes(x = x, y = y, group = measurement_id), 
            colour = "white") +
  geom_path(aes(x = x, y = y, group = measurement_id, colour = kl_hat_or - kl_hat), 
            alpha = 0.3) +
  scale_colour_gradient2(low = scales::muted("blue"), 
                         mid = "white", 
                         high = scales::muted("red"))  +
  facet_grid(cut_number(arom15arom16, 4) ~ data_type) +
  guides(colour = guide_colorbar(title = "Prediction difference")) +
  labs(x = expression("Wavenumber ["*cm^{-1}*"]"), y = "Absorbance") + 
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "Standardized normalized arom15arom16 peak height", breaks = NULL, labels = NULL)) +
  scale_x_continuous(sec.axis = sec_axis(~ . , name = "Sample type", breaks = NULL, labels = NULL)) +
  theme(
    legend.position = "bottom",
    axis.line.y.right = element_line(color = "white"),
    axis.line.x.top = element_line(color = "white")
  )
```

```{r m-prediction-differences-8}
## for holocellulose: plot prediction differences versus carb peak height conditional on different spectral variables

p_index <- with(d_pred_diff, 
                  (sample_type != "old magazines" | is.na(sample_type)) & 
                  model == "Best binned spectra" &
                  (!core_label %in% c("MB-775", "MB-930") | is.na(core_label)))
facet_n_level <- 3

p_pred_diff_hol_intensity1 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  ggplot(aes(x = intensity_3400 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point(alpha = 0.3, shape = 16) + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
  labs(y = "Prediction difference", x = expression("Intensity at 3400 cm"^{-1}), 
       title = expression("Intensity at 3400 cm"^{-1})) +
  guides(colour = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

p_pred_diff_hol_intensity2 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  ggplot(aes(x = intensity_1250 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point(alpha = 0.3, shape = 16) + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
  labs(y = "Prediction difference", x = expression("Intensity at 1250 cm"^{-1}), 
       title = expression("Intensity at 1250 cm"^{-1})) +
  guides(colour = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

p_pred_diff_hol_intensity3 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  ggplot(aes(x = intensity_1950 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point(alpha = 0.3, shape = 16) + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
  labs(y = "Prediction difference", x = expression("Intensity at 1590 cm"^{-1}), 
       title = expression("Intensity at 1590 cm"^{-1})) +
  guides(colour = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

# combine
p_pred_diff_hol_intensity1 <-
  p_pred_diff_hol_intensity1 /
  p_pred_diff_hol_intensity2 /
  p_pred_diff_hol_intensity3 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") + 
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

## same plots with Mer Bleue cores

p_index <- with(d_pred_diff, 
                  (sample_type != "old magazines" | is.na(sample_type)) & 
                  model == "Best binned spectra")

p_pred_diff_hol_intensity4 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  ggplot(aes(x = intensity_1250 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point(alpha = 0.3, shape = 16) + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
  labs(y = "Prediction difference", x = expression("Intensity at 3400 cm"^{-1}), 
       title = expression("Intensity at 3400 cm"^{-1})) +
  guides(colour = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

p_pred_diff_hol_intensity5 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  ggplot(aes(x = intensity_1250 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point(alpha = 0.3, shape = 16) + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
  labs(y = "Prediction difference", x = expression("Intensity at 1250 cm"^{-1}), 
       title = expression("Intensity at 1250 cm"^{-1})) +
  guides(colour = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

p_pred_diff_hol_intensity6 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels %>% rev())) %>%
  ggplot(aes(x = intensity_1590 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point(alpha = 0.3, shape = 16) + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut_number(carb, n = facet_n_level)) +
  labs(y = "Prediction difference", x = expression("Intensity at 1590 cm"^{-1}), 
       title = expression("Intensity at 1590 cm"^{-1})) +
  guides(colour = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_color_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )

# combine
p_pred_diff_hol_intensity2 <-
  p_pred_diff_hol_intensity4 /
  p_pred_diff_hol_intensity5 /
  p_pred_diff_hol_intensity6 + 
  plot_annotation(tag_levels = "a", tag_prefix = "(", tag_suffix = ")") + 
  plot_layout(guides = "collect") &
  theme(legend.position = "bottom")

# plots just showing boxplots of OH peak height between different data types
p_oh_height_difference <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index) %>%
  ggplot(aes(y = intensity_3400, x = data_type, fill = data_type)) +
  geom_boxplot(width = 0.5) +
  labs(y = expression(Intensity~at~3400~cm^{-1}), x = "Data type") +
  scale_fill_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  ) +
  guides(fill = guide_legend(title = "Dataset"))

```

```{r, eval=FALSE}
## Test start
p_index_test <- with(d_pred_diff, 
                  (sample_type != "old magazines"))

d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb, hol_residuals, model)
  ) %>%
  dplyr::filter(p_index_test) %>%
  ggplot(aes(x = intensity_3400 %>% scale(), y = hol_residuals, colour = model)) + 
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(model ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
  labs(y = "Prediction difference", x = expression("Intensity at 3400 cm"^{-1}), 
       title = expression("Intensity at 3400 cm"^{-1})) +
  guides(colour = guide_legend(title = "Data type")) +
  theme(legend.position = "bottom")

d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb, hol_residuals, model)
  ) %>%
  dplyr::filter(p_index_test) %>%
  ggplot(aes(x = intensity_1250 %>% scale(), y = hol_residuals, colour = model)) + 
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(model ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
  labs(y = "Prediction difference", x = expression("Intensity at 1250 cm"^{-1}), 
       title = expression("Intensity at 1250 cm"^{-1})) +
  guides(colour = guide_legend(title = "Data type")) +
  theme(legend.position = "bottom")

d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb, hol_residuals, model)
  ) %>%
  dplyr::filter(p_index_test) %>%
  ggplot(aes(x = intensity_1590 %>% scale(), y = hol_residuals, colour = model)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(model ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
  labs(y = "Prediction difference", x = expression("Intensity at 1590 cm"^{-1}), 
       title = expression("Intensity at 1590 cm"^{-1})) +
  guides(colour = guide_legend(title = "Data type")) +
  theme(legend.position = "bottom")

# same plots as originals again, but with restricted carb values

p_index_test2 <- with(d_pred_diff, 
                  (sample_type != "old magazines" | is.na(sample_type)) & 
                  model == "Best binned spectra" &
                    carb >= -1.45)

p_pred_diff_hol_intensity7 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index_test2) %>%
  ggplot(aes(x = intensity_3400 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point() + 
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
  labs(y = "Prediction difference", x = expression("Intensity at 3400 cm"^{-1}), 
       title = expression("Intensity at 3400 cm"^{-1})) +
  guides(colour = guide_legend(title = "Data type")) +
  theme(legend.position = "bottom")

p_pred_diff_hol_intensity8 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index_test2) %>%
  ggplot(aes(x = intensity_1250 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
  labs(y = "Prediction difference", x = expression("Intensity at 1250 cm"^{-1}), 
       title = expression("Intensity at 1250 cm"^{-1})) +
  guides(colour = guide_legend(title = "Data type")) +
  theme(legend.position = "bottom")

p_pred_diff_hol_intensity9 <-
  d_pred_diff_intensities %>%
  dplyr::bind_cols(
    d_pred_diff %>%
      dplyr::select(data_type, hol_hat_or, hol_hat, carb)
  ) %>%
  dplyr::filter(p_index_test2) %>%
  ggplot(aes(x = intensity_1590 %>% scale(), y = hol_hat_or - hol_hat, colour = data_type)) +
  geom_hline(yintercept = 0, colour = "grey50") +
  geom_point() +
  geom_smooth(formula = y ~ x, method = "loess", se = FALSE) +
  facet_grid(data_type ~ cut(carb, breaks = c(-1.45, -0.883, 0.454, 1.16))) +
  labs(y = "Prediction difference", x = expression("Intensity at 1590 cm"^{-1}), 
       title = expression("Intensity at 1590 cm"^{-1})) +
  guides(colour = guide_legend(title = "Data type")) +
  theme(legend.position = "bottom")
```

```{r m-prediction-differences-9}
## Klason lignin: histogram of arom15arom16 conditional on arom15arom16_rel

p_index <- with(d_pred_diff, 
                  (sample_type != "old magazines" | is.na(sample_type)) & 
                  model == "Original Gaussian") # model is in principle unimportant, this just avoids duplicates

# plot
p_pred_diff_kl_histogram1 <-
  d_pred_diff %>%
  dplyr::filter(p_index) %>%
  dplyr::mutate(data_type = factor(data_type, levels = data_type_levels)) %>%
  ggplot() +
  geom_histogram(aes(x = arom15arom16 %>% scale(), fill = data_type), bins = 30) +
  facet_wrap(~ cut_number(arom15arom16_rel, n = 4)) +
  labs(y = "Count", x = "arom15arom16") +
  guides(fill = guide_legend(title = "Dataset")) +
  theme(legend.position = "bottom") +
  scale_fill_manual(
    breaks = data_type_levels, 
    values = colors_prediction_domain
  )
```