### 2022-11-23 analysis/paper/002-reply-main.Rmd
Rework manuscript 2, update the README.

---
title: 'The need to reinterpret Hodgkins et al. (2018)'
author: "Henning Teickner$^{1, *}$, Klaus-Holger Knorr$^1$"
date: ""
output: 
  bookdown::pdf_book:
    keep_tex: true
    citation_package: natbib
    toc: false
bibliography: references
biblio-style: unsrtnat
natbiboptions: super
header-includes:
   - \usepackage{float}
   - \usepackage{lineno}
   - \linenumbers
   - \usepackage{setspace}
   - \doublespacing
fontsize: 12pt
---

```{r setup-reply-2, include=FALSE}
knitr::opts_chunk$set(fig.align = "center", 
                      fig.pos = "H",
                      echo = FALSE,
                      cache = TRUE)
```

$^1$ ILÖK, Ecohydrology and Biogeochemistry Group, University of Münster, Heisenbergstr. 2, 48149 Münster, Germany  
$^*$ Corresponding author: Henning Teickner (henning.teickner@uni-muenster.de)

@Hodgkins.2018 used spectral prediction models to estimate two major peat organic matter fractions --- holocellulose and Klason lignin --- in peat cores collected along a latitudinal gradient. Based on this, @Hodgkins.2018 argue that (sub)tropical peat deposits accumulate despite high temperatures due to higher Klason lignin over holocellulose contents, and that climate warming is a risk to northern surface peat deposits which would decompose faster due to higher holocellulose contents, but not to deeper peat which is more similar to tropical peat. In a recent study, we have shown that the spectral prediction models behind this conclusion are not valid for peat samples [@Teickner.2022f]. In addition to that, we show here that @Hodgkins.2018 did not consider important uncertainty sources in their analysis. This raises the question whether the results of @Hodgkins.2018, and of a follow-up study with more data [@Verbeke.2022], remain valid or how they have to be reinterpreted.

We show that the results of @Hodgkins.2018 are to some extent still valid for holocellulose, but only when reinterpreted in terms of differences in peak heights, which is a major obstacle to a quantitative understanding of the control of peat chemistry on decomposition. For Klason lignin, remaining uncertainties are high even after reinterpretation in terms of peak heights. Even more, there is evidence that near-surface peat Klason lignin contents and residual enrichment of Klason lignin during decomposition have been underestimated. It thus remains uncertain whether and to what extent the reported latitudinal gradient in peat chemistry will control future peatland carbon losses.

## Why are the results highly uncertain? {.unnumbered}

The original models to estimate peat holocellulose and Klason lignin contents are not valid and, highly uncertain, and thus likely biased --- especially for decomposed peat --- and therefore should not be used to estimate holocellulose and Klason lignin contents [@Teickner.2022f]. 
Besides, there are that two important sources of uncertainty have been neglected in @Hodgkins.2018: first, the prediction uncertainty of the spectral prediction models, and second, the uncertainty introduced when computing near-surface average holocellulose and Klason lignin contents from estimates for individual peat layers.  
We illustrate the impact of all these factors by recomputing the latitudinal gradient and depth profiles with the modified, probably more accurate, but still biased, models from @Teickner.2022f, while considering these uncertainty sources (supporting information S1).  
This analysis shows that both the estimated slopes for the latitudinal gradient (95% confidence intervals are [`r s1_res_holocellulose %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] and [`r s1_res_klason_lignin %>% dplyr::pull(b_latitude) %>% quantile(probs = c(0.025, 0.975)) %>% round(1) %>% paste0(collapse = ",")`] for holocellulose and Klason lignin, respectively; figure \@ref(fig:reply-res-p-sites-latitude) and supporting figure 3<!---\@ref(fig:reply-res-p-latitude-poc-slope)--->) and differences in the depth profiles are highly uncertain due to these uncertainty sources (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).

## Klason lignin contents and residual enrichment with decomposition probably are larger {.unnumbered}

In a previous publication, we have shown that the spectral prediction model for Klason lignin underestimates Klason lignin contents especially when these are $>30$ mass-% [@Teickner.2022f]. Our modified model is much less biased [@Teickner.2022f] and therefore enables a rough analysis of the order of magnitude of the bias. This allowed us to analyze what consequences the bias can have for the estimated near-surface ($\le50$ cm) peat Klason lignin contents and residual enrichment of Klason lignin during decomposition (supporting information S1).   
With our modified model, near-surface peat Klason lignin contents are on average ~10 to 15 mass-% larger across the latitudinal gradient (figure \@ref(fig:reply-res-p-sites-latitude)). Specifically, average (sub)tropical ($<45 ^\circ$N) near-surface peat Klason lignin contents are `r d_peat_pred_best_diff_mean %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude & surface_peat) %>% dplyr::pull(y_hat_diff) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude & surface_peat) %>% dplyr::pull(y_hat_diff) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-% (median, lower and upper 95% prediction interval limit) larger than with the original model. There is also tentative indication for on average larger differences between deep (sub)tropical ($<45 ^\circ$N) and high-latitude ($\ge45 ^\circ$N) peat (figure \@ref(fig:reply-res-p-depth-profiles-latitude)).    
Residual enrichment of Klason lignin during decomposition probably has been underestimated. A rough estimate for the residual enrichment of Klason lignin during decomposition is the difference in Klason lignin content between near-surface peat and deeper peat. With the modified model, this difference is on average for (sub)tropical peatlands `r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude) %>% dplyr::pull(res) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & !high_latitude) %>% dplyr::pull(res) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-% larger (figure \@ref(fig:reply-res-p-depth-profiles-latitude)). For high latitude peatlands the difference is smaller (`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & high_latitude) %>% dplyr::pull(res) %>% median() %>% magrittr::multiply_by(100) %>% round(0)` [`r d_peat_pred_best_diff_mean1 %>% dplyr::mutate(res = y_hat_diff - y_hat_or_diff) %>% dplyr::filter(variable_y == "Klason lignin" & high_latitude) %>% dplyr::pull(res) %>% quantile(probs = c(0.025, 0.975)) %>% magrittr::multiply_by(100) %>% round(0) %>% paste(collapse = ",")`] mass-%).  
Consequently, in general --- and especially in (sub)tropical peatlands --- the average residual enrichment of Klason lignin due to decomposition probably has been underestimated by the original model. A consequence of this in light of climate change is that high latitude peat deposits may be even more prone to shifts in decomposition rates than suggested by the original results.

## How to reinterpret @Hodgkins.2018 {.unnumbered}

The mid-infrared spectra (MIRS), which have been used to compute the spectral prediction models, have so far mostly been interpreted in terms qualitative concepts for organic matter fractions (e.g. @Cocozza.2003): the `carb` peak [@Hodgkins.2018] --- the only predictor for holocellulose content in @Hodgkins.2018 --- is an indicator for "carbohydrates" in mineral-poor peat [@Hodgkins.2018]; the `arom15arom16` peak [@Hodgkins.2018] --- the only predictor for Klason lignin content in @Hodgkins.2018 --- is an indicator for "aromatics" in peat with low protein and carbonate contents [@Stuart.2004; @Tatzber.2007; @Teickner.2022f] (note that we write "carbohydrates" and "aromatics" because, when interpreted only in terms of peak heights, it is remains uncertain what these organic matter (OM) fractions represent chemically or how one could extract them).

Such a qualitative interpretation of MIRS enables us to reinterpret @Hodgkins.2018, circumventing some of the uncertainties we have uncovered: All holocellulose and Klason lignin values given in @Hodgkins.2018 are directly proportional to the heights of the `carb` and `arom15arom16` peak, respectively. This removes uncertainties due to invalid spectral prediction models from the analyses. Then, we can still say that @Hodgkins.2018 show systematic latitudinal and depth-related differences in peat "carbohydrate" and "aromatics" contents.

The trade-offs are first, that we have to make the assumption that there are no latitudinal or depth-related differences in mineral, carbonate, and protein contents, and second, that we have to make the assumption that differences in heights of the two peaks are not dominated by differences in other peaks in the spectra. The reason is that differences in peak heights can be attributed to "carbohydrates" or "aromatics" only when contents of minerals or proteins and carbonates, respectively, are the same, and when changes in the area normalized height of these peaks are not caused by changes in other peaks, which can be an effect of spectral normalization [@Stuart.2004; @Tatzber.2007; @Hodgkins.2018; @Teickner.2022f].

We think that only some of these assumptions can are supported at the moment. The `carb` peak is comparatively large and has low interferences with peaks caused by other molecular structures. It should therefore be a good indicator for "carbohydrates" because mineral contents in (near-surface) peat are typically low (as validated by @Hodgkins.2018). The `arom15arom16` peak has large contributions by proteins [@Stuart.2004; @Teickner.2022f], based on peat N contents, it is not unlikely that peat protein contents differ with depth and latitude [@Sowden.1978; @Loisel.2014], and it may be more affected by other peaks since it is smaller than the `carb` peak. For this reason, we suggest that more evaluation is necessary before accepting, based on @Hodgkins.2018, that there are systematic differences in peat "aromatic" contents.

The reinterpretation in terms of "carbohydrate" and "aromatics" instead of holocellulose and Klason lignin has the clear disadvantage that we sacrifice the quantitative precision @Hodgkins.2018 aimed to attribute to the results: it is much more informative to precisely estimate a change in Klason lignin content of e.g. 5 mass-% when the average annual temperature increase by 1$^{\circ}$C than to only make qualitative statements that the relative amount of "aromatics" (without describing in detail what this OM fraction means) increases.

We think @Hodgkins.2018 have advanced a very important idea to make the interpretation of MIRS more quantitative. Only with such a quantitative approach, which links different types of measurements and the knowledge we have about them, can we develop a quantitatively precise understanding. However, we have shown that there is some more way to go. We hope that our critical analysis will help to improve this approach.

(ref:reply-res-p-sites-latitude-caption) Predicted average surface ($\le 50$ cm) peat holocellulose (a) and Klason lignin (b) contents plotted against latitude (compare with Fig. 3 in @Hodgkins.2018). Lines and shaded areas represent average predictions from regression models and 95% confidence intervals. "Modified" is the modified Bayesian hierarchical regression model which simultaneously models individual samples' contents from mid infrared spectra and the latitudinal gradient of average core near-surface peat contents. This model considers prediction uncertainty from the mid infrared spectra and from computing per-core averages. "Original" is the original linear regression model [@Hodgkins.2018] computing only the latitudinal gradient of average core near-surface contents. This model does not consider prediction uncertainty from mid infrared spectra, nor uncertainty from computing per-core averages. Points are average core near-surface contents predicted from the model ("Modified") or computed from the average predictions for individual samples ("Original") with error bars representing 95% confidence intervals. Points for "Modified" are shifted by $+0.1^\circ$.

```{r reply-res-p-sites-latitude, echo=FALSE, out.width="70%", fig.height=3.5, fig.width=6.5, fig.cap='(ref:reply-res-p-sites-latitude-caption)'}
p_poc_latitude6
```

(ref:reply-res-p-depth-profiles-latitude-caption) Predicted median holocellulose (A) and Klason lignin (B) depth profiles of peat core data classified into two latitude categories following @Hodgkins.2018 (compare with Fig. 2 in @Hodgkins.2018). Lines are averages of LOESS smoothers fitted to the predicted values by the models. Shaded regions are corresponding 95% confidence intervals, comprising prediction uncertainty in the holocellulose and Klason lignin estimates, respectively. Vertical dashed lines in each columns represent approximate surface and average mean contents as predicted using the improved models ("Modified").

```{r reply-res-p-depth-profiles-latitude, echo=FALSE, warning=FALSE, out.width="70%", fig.height=5.5, fig.width=6.5, fig.cap='(ref:reply-res-p-depth-profiles-latitude-caption)'}
p_depth_profiles_latitude
```

```{r reply-res-p-veg-peat-differnece, echo=FALSE, eval=FALSE, out.width="90%", fig.height=3.5, fig.width=7.5, fig.cap='Predicted surface ($\\le 50$ cm) peat and vegetation median holocellulose (A) and Klason lignin (B) contents for the peat cores shown in figure \\ref{fig:res-p-gaussian-beta-depth-profile-best} grouped according to \\cite{Hodgkins.2018}. Predictions of the (biased) original models are equivalent to the data shown in \\cite{Hodgkins.2018} (figure 5). Averages are median values of predictions, error bars are 95\\% prediction intervals. "Difference" is the difference between predictions for vegetation and peat.'}
p_veg_peat_difference
```

# Acknowledgements {.unnumbered}

We like to thank Jeffrey Chanton and Suzanne Hodgkins for a constructive discussion which greatly helped to improve a first version of this manuscript. This study was funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) grant no. KN 929/23-1 to Klaus-Holger Knorr and grant no. PE 1632/18-1 to Edzer Pebesma.

# Competing interests {.unnumbered}

The authors declare no competing interests.

# Author contributions {.unnumbered}

HT performed the calculations, prepared the figures, and wrote the original text. Both HT and KHK revised and edited the text. KHK provided funding for this study.

# Data and Code availability {.unnumbered}

Data and code to reproduce our analyses are available via <https://doi.org/10.5281/zenodo.6325760.>